
Extra References:
- https://arxiv.org/abs/2108.05858

We have selected a dataset from the [Econometrics Dataset Package](https://cran.r-project.org/web/packages/Ecdat/index.html) from the various datasets listed [here](https://cran.r-project.org/web/packages/Ecdat/Ecdat.pdf).

A classic example for DiD analysis is the study by Card and Krueger (1994) on the impact of minimum wage increases on employment in the fast-food industry. This study utilized data from fast-food restaurants in New Jersey and Pennsylvania.

```{r}
library(pacman)
pacman::p_load(tidyverse, fixest, did, ggplot2, plm, stargazer)
path = "D:/analysis/Econometrics/project/card_kruger/"
windows()
```

```{r}
# load the dataset
dat <- read.csv(file.path(path,'data/fast-food-data.csv'))
```

The `CardKrueger1994` dataset includes information on employment, wages, and other characteristics of fast-food restaurants before and after the **minimum wage increase in New Jersey**, with Pennsylvania serving as a control group. Familiarize yourself with the dataset by examining its structure and summary statistics

```{r}
# summary(df)
str(dat)
```

### Cleaning

```{r}
long_df <- dat %>%
  select(-ends_with("2")) %>%  # Drop A2, B2, C2 from original dataset
  rename(store_id = X) %>%
  mutate(post = 0) %>% 
  bind_rows(
    dat %>%
      rename(store_id = X) %>%
      select(store_id,chain,state,ends_with("2")) %>%
      rename_with(~ gsub("2$", "", .x)) %>% 
      mutate(post = 1)
  )
df <- long_df %>% mutate(emp_total = empft + (emppt/2))
    select(store_id, state, post, empft, emppt, nmgrs, wage_st, chain , co_owned) 
str(df) 
```

```{r}
summary(df)
```

### Defining Treatment and Control Groups

```{r}
```


### Compare means before and after Policy Change

```{r}
df %>%
  group_by(state, post) %>%
  summarise(avg_emp = mean(emp_total, na.rm = TRUE))

```


```{r}
df %>%
  group_by(state, post) %>%
  summarise(avg_emp = mean(emp_total, na.rm = TRUE)) %>%
  ggplot(aes(x = post, y = avg_emp, color = as.factor(state))) +
  geom_line() + geom_point() +
  labs(title = "Average Employment Before and After Minimum Wage Change",
       x = "Pre (0) vs Post (1)",
       y = "Average Employment",
       color = "State") +
  theme_minimal()

```


## Running the DiD Model

```{r}

```