
Extra References:
- https://arxiv.org/abs/2108.05858

We have selected a dataset from the [Econometrics Dataset Package](https://cran.r-project.org/web/packages/Ecdat/index.html) from the various datasets listed [here](https://cran.r-project.org/web/packages/Ecdat/Ecdat.pdf).

A classic example for DiD analysis is the study by Card and Krueger (1994) on the impact of minimum wage increases on employment in the fast-food industry. This study utilized data from fast-food restaurants in New Jersey and Pennsylvania.

```{r}
library(pacman)
pacman::p_load(tidyverse, fixest, did, ggplot2, plm, stargazer)
path = "D:/analysis/Econometrics/project/card_kruger/"
windows()
```

```{r}
# load the dataset
df <- read.csv(file.path(path,'data/fast-food-data.csv'))
```

The `CardKrueger1994` dataset includes information on employment, wages, and other characteristics of fast-food restaurants before and after the **minimum wage increase in New Jersey**, with Pennsylvania serving as a control group. Familiarize yourself with the dataset by examining its structure and summary statistics

```{r}
# summary(df)
str(df)
```

### Defining Treatment and Control Groups

```{r}
df <- df %>%
  mutate(
    treatment = ifelse(state == 1, 1, 0),  # 1 if NJ (treatment), 0 if PA (control)
    post_treatment = ifelse(!is.na(empft2), 1, 0),  # 1 if second interview data exists
    emp_total1 = empft + (emppt / 2),  # Convert part-time to full-time equivalents
    emp_total2 = empft2 + (emppt2 / 2)  # After minimum wage increase
)

```


### Compare means before and after Policy Change

```{r}
df %>%
  group_by(treatment, post_treatment) %>%
  summarise(
    mean_emp = mean(emp_total1, na.rm = TRUE),
    mean_emp_after = mean(emp_total2, na.rm = TRUE)
  )

```


```{r}
df %>%
  group_by(state, post_treatment) %>%
  summarise(avg_emp = mean(emp_total1, na.rm = TRUE)) %>%
  ggplot(aes(x = post_treatment, y = avg_emp, color = as.factor(state))) +
  geom_line() + geom_point() +
  labs(title = "Average Employment Before and After Minimum Wage Change",
       x = "Pre (0) vs Post (1)",
       y = "Average Employment",
       color = "State") +
  theme_minimal()

```


## Running the DiD Model

```{r}
did_model <- feols(emp_total2 ~ treatment * post_treatment, data = df)
summary(did_model)
```